### 2021-04-21 paper.tex
Report

\def\paperversiondraft{draft}
\def\paperversionblind{blind}
\def\paperversioncamera{camera}

% If no special paper-version is requested, compile in draft mode
\ifx\paperversion\paperversionblind
\else
  \ifx\paperversion\paperversioncamera
  \else
     \def\paperversion{draft}
  \fi
\fi

\def\grammarlyon{on}

\ifx\grammarly\grammarlyon
\def\review{}
\else
\def\review{review,}
\fi

\ifx\paperversion\paperversiondraft
  \documentclass[11pt, manuscript,\review anonymous]{acmart}
\fi

\ifx\paperversion\paperversionblind
  \documentclass[manuscript,\review anonymous]{acmart}
\fi

\ifx\paperversion\paperversioncamera
  \documentclass[11pt, manuscript, nonacm, natbib, oneside]{acmart}\settopmatter{}
\fi

\geometry{twoside=false}
\usepackage{colortbl}

% 'draftonly' environment
\usepackage{environ}
\ifx\paperversion\paperversiondraft
\newenvironment{draftonly}{}{}
\else
\NewEnviron{draftonly}{}
\fi

% Most PL conferences are edited by conference-publishing.com. Follow their
% advice to add the following packages.
%
% The first enables the use of UTF-8 as character encoding, which is the
% standard nowadays. The second ensures the use of font encodings that support
% accented characters etc. (Why should I use this?). The mictotype package
% enables certain features 'to­wards ty­po­graph­i­cal per­fec­tion
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{microtype}

\usepackage{xargs}
\usepackage{lipsum}
\usepackage[textsize=tiny]{todonotes}
\usepackage{xparse}
\usepackage{xifthen, xstring}
\usepackage[normalem]{ulem}
\usepackage{xspace}
\usepackage{marginnote}
\makeatletter
\font\uwavefont=lasyb10 scaled 652

\newcommand\colorwave[1][blue]{\bgroup\markoverwith{\lower3\p@\hbox{\uwavefont\textcolor{#1}{\char58}}}\ULon}
\makeatother

\ifx\paperversion\paperversiondraft
\newcommand\createtodoauthor[2]{%
\def\tmpdefault{emptystring}
\expandafter\newcommand\csname #1\endcsname[2][\tmpdefault]{\def\tmp{##1}\ifthenelse{\equal{\tmp}{\tmpdefault}}
   {\todo[linecolor=#2,backgroundcolor=#2,bordercolor=#2]{\textbf{#1:} ##2}}
   {\ifthenelse{\equal{##2}{}}{\colorwave[#2]{##1}\xspace}{\todo[linecolor=#2,backgroundcolor=#2,bordercolor=#2]{\textbf{#1:} ##2}\colorwave[#2]{##1}}}}
\expandafter\newcommand\csname #1f\endcsname[2][\tmpdefault]{
	\marginnote{
		\todo[inline,linecolor=#2,backgroundcolor=#2,bordercolor=#2]{\textbf{#1 (Figure):} ##2}}
   }
}

\else
\newcommand\createtodoauthor[2]{%
\def\tmpdefault{emptystring}
\expandafter\newcommand\csname #1\endcsname[2][\tmpdefault]{\def\tmp{##1}\ifthenelse{\equal{\tmp}{\tmpdefault}}
   {}
   {\ifthenelse{\equal{##2}{}}{##1\xspace}{##1}}}
\expandafter\newcommand\csname #1f\endcsname[2][\tmpdefault]{
   }
   }
\fi

% Broaden margins to make room for todo notes
\makeatletter
\patchcmd{\@addmarginpar}{\ifodd\c@page}{\ifodd\c@page\@tempcnta\m@ne}{}{}
\makeatother
\ifx\paperversion\paperversiondraft
  \makeatletter
  \if@ACM@journal
    \geometry{asymmetric}
    \paperwidth=\dimexpr \paperwidth + 3.5cm\relax
    \oddsidemargin=\dimexpr\oddsidemargin + 0cm\relax
    \evensidemargin=\dimexpr\evensidemargin + 0cm\relax
    \marginparwidth=\dimexpr \marginparwidth + 3cm\relax
    \setlength{\marginparwidth}{4.6cm}
    % This makeatletter box helps to move notes to the right
    \makeatletter
    \long\def\@mn@@@marginnote[#1]#2[#3]{%
      \begingroup
        \ifmmode\mn@strut\let\@tempa\mn@vadjust\else
          \if@inlabel\leavevmode\fi
          \ifhmode\mn@strut\let\@tempa\mn@vadjust\else\let\@tempa\mn@vlap\fi
        \fi
        \@tempa{%
          \vbox to\z@{%
            \vss
            \@mn@margintest
            \if@reversemargin\if@tempswa
                \@tempswafalse
              \else
                \@tempswatrue
            \fi\fi
            %\if@tempswa
              \rlap{%
                \if@mn@verbose
                  \PackageInfo{marginnote}{xpos seems to be \@mn@currxpos}%
                \fi
                \begingroup
                  \ifx\@mn@currxpos\relax\else\ifx\@mn@currxpos\@empty\else
                      \kern-\dimexpr\@mn@currxpos\relax
                  \fi\fi
                  \ifx\@mn@currpage\relax
                    \let\@mn@currpage\@ne
                  \fi
                  \if@twoside\ifodd\@mn@currpage\relax
                      \kern\oddsidemargin
                    \else
                      \kern\evensidemargin
                    \fi
                  \else
                    \kern\oddsidemargin
                  \fi
                  \kern 1in
                \endgroup
                \kern\marginnotetextwidth\kern\marginparsep
                \vbox to\z@{\kern\marginnotevadjust\kern #3
                  \vbox to\z@{%
                    \hsize\marginparwidth
                    \linewidth\hsize
                    \kern-\parskip
                    \marginfont\raggedrightmarginnote\strut\hspace{\z@}%
                    \ignorespaces#2\endgraf
                    \vss}%
                  \vss}%
              }%
          }%
        }%
      \endgroup
    }
    \makeatother
  \else
    \paperwidth=\dimexpr \paperwidth + 6cm\relax
    \oddsidemargin=\dimexpr\oddsidemargin + 3cm\relax
    \evensidemargin=\dimexpr\evensidemargin + 3cm\relax
    \marginparwidth=\dimexpr \marginparwidth + 3cm\relax
    \setlength{\marginparwidth}{4.6cm}
  \fi
  \makeatother
\fi

% We use the following color scheme
% 
% This scheme is both print-friendly and colorblind safe for
% up to four colors (including the red tones makes it not
% colorblind safe any more)
%
% https://colorbrewer2.org/#type=qualitative&scheme=Paired&n=4

\definecolor{pairedNegOneLightGray}{HTML}{cacaca}
\definecolor{pairedNegTwoDarkGray}{HTML}{827b7b}
\definecolor{pairedOneLightBlue}{HTML}{a6cee3}
\definecolor{pairedTwoDarkBlue}{HTML}{1f78b4}
\definecolor{pairedThreeLightGreen}{HTML}{b2df8a}
\definecolor{pairedFourDarkGreen}{HTML}{33a02c}
\definecolor{pairedFiveLightRed}{HTML}{fb9a99}
\definecolor{pairedSixDarkRed}{HTML}{e31a1c}

\createtodoauthor{grosser}{pairedOneLightBlue}
\createtodoauthor{authorTwo}{pairedTwoDarkBlue}
\createtodoauthor{authorThree}{pairedThreeLightGreen}
\createtodoauthor{authorFour}{pairedFourDarkGreen}
\createtodoauthor{authorFive}{pairedFiveLightRed}
\createtodoauthor{authorSix}{pairedSixDarkRed}

\graphicspath{{./images/}}

% Define macros that are used in this paper
%
% We require all macros to end with a delimiter (by default {}) to enusure
% that LaTeX adds whitespace correctly.
\makeatletter
\newcommand\requiredelimiter[2][########]{%
  \ifdefined#2%
    \def\@temp{\def#2#1}%
    \expandafter\@temp\expandafter{#2}%
  \else
    \@latex@error{\noexpand#2undefined}\@ehc
  \fi
}
\@onlypreamble\requiredelimiter
\makeatother

\newcommand\newdelimitedcommand[2]{
\expandafter\newcommand\csname #1\endcsname{#2}
\expandafter\requiredelimiter
\csname #1 \endcsname
}

\newdelimitedcommand{toolname}{Tool}

% Print \autoref as "Section X.Y.Z"
\renewcommand*{\sectionautorefname}{Section}
\renewcommand*{\subsectionautorefname}{Section}
\renewcommand*{\subsubsectionautorefname}{Section}

% \circled command to print a colored circle.
% \circled{1} pretty-prints "(1)"
% This is useful to refer to labels that are embedded within figures.
\usepackage{tikz}
\usetikzlibrary{arrows}
\usetikzlibrary{shapes}
\newcommand*\circled[1]{\tikz[baseline=(char.base)]{
            \node[shape=circle,fill=pairedOneLightBlue,inner sep=1pt] (char) {#1};}}

%% Note: Authors migrating a paper from traditional SIGPLAN
%% proceedings format to PACMPL format should change 'sigplan' to
%% 'acmsmall'.

%% Some recommended packages.
\usepackage{booktabs}   %% For formal tables:
                        %% http://ctan.org/pkg/booktabs
\usepackage{subcaption} %% For complex figures with subfigures/subcaptions
                        %% http://ctan.org/pkg/subcaption
\usepackage{bibentry}
\nobibliography*

\makeatletter\if@ACM@journal\makeatother
%% Journal information (used by PACMPL format)
%% Supplied to authors by publisher for camera-ready submission
\acmJournal{PACMPL}
\acmVolume{1}
\acmNumber{1}
\acmArticle{1}
\acmYear{2017}
\acmMonth{1}
\acmDOI{10.1145/nnnnnnn.nnnnnnn}
\startPage{1}
\else\makeatother
%% Conference information (used by SIGPLAN proceedings format)
%% Supplied to authors by publisher for camera-ready submission
\acmConference[PL'17]{ACM SIGPLAN Conference on Programming Languages}{January 01--03, 2017}{New York, NY, USA}
\acmYear{2017}
\acmISBN{978-x-xxxx-xxxx-x/YY/MM}
\acmDOI{10.1145/nnnnnnn.nnnnnnn}
\startPage{1}
\fi

%% Copyright information
%% Supplied to authors (based on authors' rights management selection;
%% see authors.acm.org) by publisher for camera-ready submission
\setcopyright{none}             %% For review submission
%\setcopyright{acmcopyright}
%\setcopyright{acmlicensed}
%\setcopyright{rightsretained}
%\copyrightyear{2017}           %% If different from \acmYear

%% Bibliography style
%% Citation style
%% Note: author/year citations are required for papers published as an
%% issue of PACMPL.
%\citestyle{acmauthoryear}  %% For author/year citations
%\citestyle{acmnumeric}     %% For numeric citations
%\setcitestyle{nosort}      %% With 'acmnumeric', to disable automatic
                            %% sorting of references within a single citation;
                            %% e.g., \cite{Smith99,Carpenter05,Baker12}
                            %% rendered as [14,5,2] rather than [2,5,14].
%\setcitesyle{nocompress}   %% With 'acmnumeric', to disable automatic
                            %% compression of sequential references within a
                            %% single citation;
                            %% e.g., \cite{Baker12,Baker14,Baker16}
                            %% rendered as [2,3,4] rather than [2-4].

\begin{document}

%% Title information
\title[]{
	{
	\centering
	Scientific Report - Summary of Research\\
	~\\
Automatized compilation of sequential software to a diverse set of
hardware accelerators\\
}
~\\
\vspace{18em}
\large
Grant number: PZ00P2 168016\\
Period: 01.05.2018– 31.12.2020\\
Grantee: Tobias Grosser\\
\today
}         %% [Short Title] is optional;
                                        %% when present, will be used in
                                        %% header instead of Full Title.

%% Author information
%% Contents and number of authors suppressed with 'anonymous'.
%% Each author should be introduced by \author, followed by
%% \authornote (optional), \orcid (optional), \affiliation, and
%% \email.
%% An author may have multiple affiliations and/or emails; repeat the
%% appropriate command.
%% Many elements are not rendered, but should be provided for metadata
%% extraction tools.

%% Author with single affiliation.

% An abstract should consist of six main sentences:
%  1. Introduction. In one sentence, what’s the topic?
%  2. State the problem you tackle.
%  3. Summarize (in one sentence) why nobody else has adequately answered the research question yet.
%  4. Explain, in one sentence, how you tackled the research question.
%  5. In one sentence, how did you go about doing the research that follows from your big idea.
%  6. As a single sentence, what’s the key impact of your research?

% (http://www.easterbrook.ca/steve/2010/01/how-to-write-a-scientific-abstract-in-six-easy-steps/)

% Only add ACM notes and keywords in camera ready version
% Drop citations and footnotes in draft and blind mode.
\ifx\paperversion\paperversioncamera
%% 2012 ACM Computing Classification System (CSS) concepts
%% Generate at 'http://dl.acm.org/ccs/ccs.cfm'.
%% End of generated code

%% Keywords
%% comma separated list

\settopmatter{printacmref=false} % Removes citation information below abstract
\renewcommand\footnotetextcopyrightpermission[1]{} % removes footnote with conference information in first column
\fi

%% \maketitle
%% Note: \maketitle command must come after title commands, author
%% commands, abstract environment, Computing Classification System
%% environment and commands, and keywords command.
\maketitle
\newpage
\ifx\grammarly\grammarlyon 
\onecolumn 
\else 
\fi

\section{Research Work Conducted}

The primary goals of this Ambizione research project were proposed as:
\begin{itemize}
\item Develop a multi-target heterogeneous compute compiler for C++ and Julia code that generates binaries which effectively exploit
OpenCL-capable accelerators including GPU, Xeon-Phi, and FPGA.
\item Research new techniques to scale polyhedral optimization to large kernels and to improve their efficiency to a level that permits
the effective exploration of a large set of mapping choices.
\item Derive effective performance models based on hybrid analytical/machine-learning techniques and use them to analyze and optimize accelerator mapping decisions in a large body of existing software.
\item Evolve our new performance modeling techniques to work within the resource constraints of Julias dynamic run-time system and use Julia to experiment with distributed learning techniques.
\end{itemize}

The focus of year three in this Ambizione project was the integration of a scalable
polyhedral core infrastructure into real-world compilers (e.g, LLVM/MLIR),
establishing backends for targetting accelerators such as GPUs and FPGAs, and,
finally, developing good performance models. As mentioned in the year 2
report, our focus moved from Julia to MLIR as a compiler and programming
language infrastructure.

In the following, we discuss four milestones we identified as important to
reach these goals:

\begin{enumerate}
\item Fast, robust, and correct polyhedral libraries
\item Automatic Raising to high-level descriptions
\item Fast GPU code generation demonstrated via a climate DSL
\item Code generation for emerging architectures and FPGA
\item Scalable performance analysis within a compiler.
\end{enumerate}

\subsection{Fast, robust, and correct polyhedral libraries}
Presburger arithmetic provides the mathematical core for the polyhedral
compilation techniques that drive analytical cache models, loop optimization
for ML and HPC, formal verification, and even hardware design.  Polyhedral
compilation is widely regarded as being slow due to the potentially high
computational cost of the underlying Presburger libraries. Researchers
typically use these libraries as powerful black-box tools, but a lack of
internal documentation and decade-old C implementations hold back broader
performance-optimization efforts. With FPL, we introduced a new library for
Presburger arithmetic for the MLIR compiler built from the ground up in modern
C++. We carefully documented its internal algorithmic foundations, use
lightweight C++ data structures to minimize memory management costs, and deployed
transprecision computing across the entire library to effectively exploit
machine integers and vector instructions. On a newly-developed comprehensive
benchmark suite for Presburger arithmetic, we show a 5.25x speedup in total
runtime over the state-of-the-art library isl in its default configuration and
3.14x when over a variant of isl optimized with element-wise transprecision
computing. We expect that the availability of a well-documented and fast
Presburger library will accelerate the adoption of polyhedral compilation
techniques in production compilers.

The central components of this library have already been upstreamed into the
open-source LLVM/MLIR project and are usable by the wider research and 
industrial community. The work in this subproject has been published in \citet{grosser2020fast} and
is under review in \citet{pichanatan2021}.

\subsection{Automatic Raising to high-level descriptions}

Multi-level intermediate representations (IR) show great promise for lowering
the design costs for domain-specific compilers by providing a reusable,
extensible, and non-opinionated framework for expressing domain-specific and
high-level abstractions directly in the IR. But, while such frameworks support
the progressive lowering of high-level representations to low-level IR, they do
not raise in the opposite direction. Thus, the entry point into the compilation
pipeline defines the highest level of abstraction for all subsequent
transformations, limiting the set of applicable optimizations, in particular
for general-purpose languages that are not semantically rich enough to model
the required abstractions.  We propose Progressive Raising, a complementary
approach to the progressive lowering in multi-level IRs that raises from lower
to higher-level abstractions to leverage domain-specific transformations for
low-level representations. We further introduce MultiLevel Tactics, our
declarative approach for progressive raising, implemented on top of the MLIR
framework, and demonstrate the progressive raising from affine loop nests
specified in a general-purpose language to high-level linear algebra
operations.  Our raising paths leverage subsequent high-level domain-specific
transformations with significant performance improvements.

All this work has been has already been upstreamed into the open-source
LLVM/MLIR project and is usable by the wider research and industrial community.
The work in this subproject has been published in \citet{chelini2021mlir}.

\subsection{Fast GPU code generation demonstrated via a climate DSL}

Most compilers have a single core intermediate representation (IR) (e.g., LLVM)
sometimes complemented with vaguely-defined IR-like data structures. This IR is
commonly low-level and close to machine instructions. As a result,
optimizations relying on domain-specific information are either not possible or
require complex analysis to recover the missing information. In contrast,
multi-level rewriting instantiates a hierarchy of dialects (IRs), lowers
programs level-by-level, and performs code transformations at the most suitable
level. We demonstrate the effectiveness of this approach for the weather and
climate domain. In particular, we develop a prototype compiler and design
stencil- and GPU-specific dialects based on a set of newly introduced design
principles. We find that two domain-specific optimizations (500 lines of code)
realized on top of LLVM’s extensible MLIR compiler infrastructure suffice to
outperform state-of-the-art solutions. In essence, multi-level rewriting
promises to herald the age of specialized compilers composed from domain- and
target-specific dialects implemented on top of a shared infrastructure.

The GPU backend of the open-earth compiler has already been upstreamed into
the open-source LLVM/MLIR project and is useable by the wider research
and industrial community. The work in this subproject has been published in \citet{gysi2020ooc}

\subsection{Code Generation for Emerging Architectures and FPGA}

Automatic optimization for emerging architectures and FPGA has been a principal
objective of this work. We exploited the polyhedral techniques we developed in
earlier tasks to generate high-performance code for emerging memory
architectures such as racetrack memories to optimize polyhedral loop nests and
deep neural networks. We also proposed infrastructure for an open-source
hardware stack that allows for the compilation and simulation of System Verilog
code. We also observed an increased use of systems with where the data models
(e.g., pointer sizes) differ between host systems and accelerators. To ensure
the automatic code generation for such architectures that also allows for
automatic data movement, we expanded LLVM and its OpenMP software stack to
support mixed-data-model offloading.

The LLHD compiler has been upstreamed as open-source into the LLVM/CIRCT
project. The work in this subproject has been published in \citet{khan2020polyhedral},
\citet{schueki2020llhd}, \citet{kurth2020mixed}, and \citet{kourtis2020compiling}.

\subsection{Scalable performance analysis within a compiler}

Performance models are well-known instruments to understand the scaling
behavior of parallel applications. They express how performance changes as key
execution parameters, such as the number of processes or the size of the input
problem, vary. Besides reasoning about program behavior, such models can also
be automatically derived from performance data. This is called empirical
performance modeling. While this sounds simple at the first glance, this
approach faces several serious interrelated challenges, including expensive
performance measurements, inaccuracies inflicted by noisy benchmark data, and
overall complex experiment design, starting with the selection of the right
parameters. The more parameters one considers, the more experiments are needed
and the stronger the impact of noise. In this paper, we show how taint
analysis, a technique borrowed from the domain of computer security, can
substantially improve the modeling process, lowering its cost, improving model
quality, and help validate performance models and experimental setups.

In a subsequent step, we analyzed the optimality of function inlining as a means
to minimize code size and alter expose further optimization opportunities. This
work has shown promising code size reductions on SPEC 2017 and the LLVM compiler.
We hope to use it in the future as an enabling transformation for large-scale
automatic accelerator mapping.

The work in this subproject has been published in \citet{copic2021} and is under review in
\citet{theo2021}.

\section{Deviations}

\section{Important Events}

\section{Contributions}

\section{Research Output}

\subsection{Publications (published)}
\begin{enumerate}
	\item \bibentry{chelini2021mlir}
	\item \bibentry{10.1145/3437801.3441613}
	\item \bibentry{grosser2020fast}
	\item \bibentry{chelini2020automatic}
	\item \bibentry{khan2020polyhedral}
	\item \bibentry{schueki2020llhd}
	\item \bibentry{kurth2020mixed}
	\item \bibentry{kourtis2020compiling}
\end{enumerate}
\subsection{Publications (under review)}
\begin{enumerate}
	\item \bibentry{gysi2020ooc}
	\item \bibentry{fehr2021}
	\item \bibentry{anurudth2021}
	\item \bibentry{theo2021}
	\item \bibentry{pichanatan2021}
	\item \bibentry{copic2021}
\end{enumerate}

%% Bibliography
\bibliographystyle{plainnat}
\nobibliography{references}

\end{document}